---
title: 'Data Product :: Shiny App :: BoxPlot Demo'
author: "J R Shrestha"
date: "May 31, 2016"
output: slidy_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```


#Introduction

This is the documentation for the data product powered by Shiny App for the Assignment project of Coursera Data Science course.

It generates a dataframe having two fields: gender and salary with random values and produces a boxplot for salary against each gender. The user can select total number of samples, mean, standard deviation for randomization of values along with some bias.

## Application URLs

Application published at:

https://jr-shrestha.shinyapps.io/DataProduct/

Code can be obtained from:

https://github.com/jr-shrestha/DataProduct

#Functionality

- When the application is launched, a box plot is produced with default values 

- User changes any of the following values through the sliders and the box plot is updated as per the user input:

    * Total Number of samples: total number of random observations to generate
    * Mean Salary: mean of the random salary values being generated (input for rnorm)
    * Standard Deviation: standard deviation of the random salary values being generated (input for rnorm)
    * Bias (in %): just to increment the salary of a gender by certain % and do the opposite to the other gender
        
        - +ve value will increse the salaries of male gender and decrease the salaries of female gender
        - -ve value will do jus the opposite

#Implementation of server.R

```{r, echo=TRUE, eval=FALSE}
library(UsingR)
shinyServer(
  function(input, output) {
    output$newBoxPlot <- renderPlot({
      #Extract the variables from UI into local variables 
      totalNum <- input$totalNum
      meanVal  <- input$meanVal
      sdevVal  <- input$sdevVal
      biasVal  <- input$biasVal
      
      #Construct a dataframe of gender and salary with random values
      gender <- as.factor(sample(c('male', 'female'), totalNum, replace=TRUE))
      salary <- rnorm(totalNum, mean=meanVal, sd=sdevVal)
      dFrame <- data.frame(gender,salary) 
      
      #Introduce some bias in the salary as per the gender
      dFrame[which(gender=='female'),]$salary <- dFrame[which(gender=='female'),]$salary * (1-biasVal/100)
      dFrame[which(gender=='male'),]$salary <- dFrame[which(gender=='male'),]$salary * (1+biasVal/100)
      
      boxplot(salary~gender, data = dFrame, xlab='Gender', ylab='salary')
    })
  }
)
```

#Implementation of ui.R

```{r, echo=TRUE, eval=FALSE}
shinyUI(pageWithSidebar(
  headerPanel("BoxPlot Demo: Monthly Salary Vs Income"),
  sidebarPanel(
    h3('Input for randomization'),
    sliderInput('totalNum', 'Total Number of Samples', value = 500, min = 100, max = 1000, step = 50),
    sliderInput('meanVal', 'Mean Salary', value = 50000, min = 10000, max = 90000, step = 1000),
    sliderInput('sdevVal', 'Standard Deviation', value = 10000, min = 0, max = 20000, step = 1000),
    
    h3('Introduce Bias'),
    p('+ve value --> increase salary in male'),
    p('-ve value --> increase salary in female'),
    sliderInput('biasVal', 'Bias (in %)', value = 0, min = -50, max = 50, step = 5)
  ),
  
  mainPanel(
    plotOutput('newBoxPlot')
  )
))
```
##Thank you!


